### 2024-09-26 conclusion.tex
new

\chapter{Conclusion} \label{chap:conclusion}

In this thesis, we have discussed the use of nested simulation procedures for risk management applications of financial derivatives and insurance products.