### 2025-01-07 pre_private/tikz/illustration_mpns.tex
new

\documentclass[tikz]{standalone}
\usepackage{tikz}
\usepackage{amssymb}

\usetikzlibrary{shapes,arrows,positioning, calc, fit, decorations.pathreplacing}

\begin{document}

\begin{tikzpicture}[scale=1]
        \draw[-latex] (0,0)--(10,0);
        \draw[-latex] (0,0)--(-1,1.5) node[above] {$\blacksquare$};
        \draw[-latex] (0,0)--(0,1.5) node[above] {$\blacksquare$};
        \draw[-latex] (0,0)--(1,1.5) node[above] {$\blacksquare$};
        \node at (0.5,1.2) {\small$\dots$};
        \node at (0.5,1.85) {\small$\dots$};

        \draw[-latex] (3,0)--(2,1.5) node[above] {$\blacksquare$};
        \draw[-latex] (3,0)--(3,1.5) node[above] {$\blacksquare$};
        \draw[-latex] (3,0)--(4,1.5) node[above] {$\blacksquare$};
        \node at (3.5,1.2) {\small$\dots$};
        \node at (3.5,1.85) {\small$\dots$};

        \draw[-latex] (6,0)--(5,1.5) node[above] {$\blacksquare$};
        \draw[-latex] (6,0)--(6,1.5) node[above] {$\blacksquare$};
        \draw[-latex] (6,0)--(7,1.5) node[above] {$\blacksquare$};
        \node at (6.5,1.2) {\small$\dots$};
        \node at (6.5,1.85) {\small$\dots$};

        \draw[-latex] (9,0)--(8,1.5) node[above] {$\blacksquare$};
        \draw[-latex] (9,0)--(9,1.5) node[above] {$\blacksquare$};
        \draw[-latex] (9,0)--(10,1.5) node[above] {$\blacksquare$};
        \node at (9.5,1.2) {\small$\dots$};
        \node at (9.5,1.85) {\small$\dots$};

        \node at (7.5,0.7) {\Huge $\dots$};

        \draw [decorate,  decoration = {brace,    raise=1pt,    amplitude=5pt}] (-1,2) --  (1,2);
        \draw [decorate,  decoration = {brace,    raise=1pt,    amplitude=5pt}] (2,2) --  (4,2);
        \draw [decorate,  decoration = {brace,    raise=1pt,    amplitude=5pt}] (5,2) --  (7,2);
        \draw [decorate,  decoration = {brace,    raise=1pt,    amplitude=5pt}] (8,2) --  (10,2);
        \draw [decorate,  decoration = {brace,    raise=1pt,    amplitude=12pt}] (0,2.4) --  (9,2.4);
        \node at (4.5,-0.5) {$S_i$};
        \node at (4.5,3.5) {$\hat L_i$};
        \node at (4.5,-1.3) {1 Outer Path for VAs};
\end{tikzpicture}

\end{document}